---
title: "Final Exam - Estimating Views"
author: "Vishal Gupta"
date: "5/1/2019"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=3,fig.align = "center")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(randomForest)
library(rpart)
library(rpart.plot)
library(caret)
library(e1071)
```

#### Problem 1 - Exploratory Data Analysis

Load youtube_train.csv into a data frame called train.
```{r}
yt_train = read.csv("~/Downloads/youtube_train.csv")
yt_test = read.csv("~/Downloads/youtube_test.csv")
```

How many rows are in the training dataset?

```{r}
yt_train %>% nrow()
```
 
What is the average amount of likes per video in the training dataset?

```{r}
yt_train %>% summarise(ct = mean(likes))
```
 
What is the category of the video with most views in the training set?

```{r}
yt_train %>% group_by(category) %>% summarise(ct = sum(as.numeric(views))) %>% arrange(desc(ct))
```

- **Music**
- Sports
- Comedy
- Entertainment
- Shows
- Film and Animation

Which category has the least amount of dislikes in the training set?

```{r}
yt_train %>% group_by(category) %>% summarise(ct = sum(dislikes)) %>% arrange(ct)
```

- Education
- Pets and Animals
- **Shows**
- People and Blogs
- Comedy
- Science and Technology
- Sports

In the training set, out of the videos with at least 1 million likes, how many have at least 100,000 comments?
```{r}
yt_train %>% filter(likes>=1000000) %>% filter(comment_count>=100000) %>% nrow() 
```


#### Problem  2.1 - Simple Linear Regression

For the rest of this problem, we will be working with log(views), log(likes), log(dislikes), and log(comment_count), which helps us manage the outliers with excessively large amounts of views, likes, dislikes, and comments. The values of log(views), log(likes), log(dislikes), and log(comment_count) are found in the columns logviews, loglikes, logdislikes and logcomments, respectively.

What is the value of log(views) that our baseline model predicts?

```{r}
mean(log(yt_train$views))
```
 
What is the correlation between log(views) and log(dislikes) in the training set?

```{r}
cor(log(yt_train$views),log(yt_train$dislikes))
```

 
Choose the most reasonable answer from the following statements:

- Higher log of dislikes are associeted with higher log of views, likely because the popular videos often have many dislikes.
- High log of dislikes are associated with less log of views.
- There is no association between log of dislikes and log of views.

#### Problem  2.2 - Simple Linear Regression (cont'd)

Load youtube_test.csv into a data frame called test.
Create a linear model that predicts log(views) using log(dislikes).

```{r}
yt_model = lm(logviews ~ logdislikes,data = yt_train)
```

What is the coefficient of log(dislikes)?
```{r}
summary(yt_model)
```

What is the R2 on the test set?
```{r}
y_pred = predict(yt_model,yt_test,type = 'response')
y_true = yt_test$logviews
rss = sum((y_pred - y_true) ^ 2)
tss = sum((y_true - mean(y_true)) ^ 2)
1 - rss/tss
```


#### Problem  3 - Adding More Variables

As good practice, it is always helpful to first check for multicollinearity before running larger models.

Examine the correlation between the following variables:

logdislikes, loglikes, logcomments, tags, publish_month, and trending_month
```{r}
cor(yt_train$loglikes,yt_train$logdislikes)
cor(yt_train$logcomments,yt_train$logdislikes)
cor(yt_train$tags,yt_train$logcomments)
cor(yt_train$trending_month,yt_train$tags)
cor(yt_train$publish_month,yt_train$trending_month)
cor(yt_train$logcomments,yt_train$loglikes)
```

Which of the following pairs of variables have correlation with magnitude above 0.8? Select all that apply.


- **logdislikes**, **loglikes**
- **logcomments**, **logdislikes**
- tags, logcomments
- trending_month, tags
- **publish_month, trending_month**
- **logcomments, loglikes**

Create a linear model that predicts log(views) using the following variables:

logdislikes, tags, and trending_month.

```{r}
yt_model_2 = lm(logviews ~ logdislikes + tags + trending_month,data = yt_train)
```


We have excluded loglikes, logcomments, and publish_month due to concerns about multicollinearity.

What is the value of the intercept?
```{r}
summary(yt_model_2)
```

 
What is the R2 on the test set?
```{r}
y_pred = predict(yt_model_2,yt_test,type = 'response')
y_true = yt_test$logviews
rss = sum((y_pred - y_true) ^ 2)
tss = sum((y_true - mean(y_true)) ^ 2)
1 - rss/tss
```


#### Problem  4 - Interpreting Linear Regression

Using the model from Problem 3, which of the following variables are significant at a level of 0.001 (p-value below 0.001)? Select all that apply.
```{r}
summary(yt_model_2)
```


- **logdislikes**
- loglikes
- **tags**
- logviews
- **trending_month**

Using the model from Problem 3, how would you interpret the coefficient of tags?

- **All else being equal, an increase in tags is associated with a 1.655e-04 increase in log(views).**
- All else being equal, an increase in tags is associated with a 1.655e-04 decrease in log(views).

Using the simple model from Problem 2, if the amount of dislikes is 1000, how many views does the model predict the video has?

- 795.2351
- 3182852
- 1906.064
- **928263.8**


#### Problem  5 - CART and Random Forest

In addition to the logistic regression model, we can also train a regression tree. Use the same variable as used in the simple model, logdislikes. Train a regression tree with cp = 0.05.
```{r}
yt_model_cart = rpart(logviews ~ logdislikes,yt_train, cp=0.05)
```

Looking at the plot of the tree, how many different predicted values are there?
```{r}
prp(yt_model_cart)
```
 
What is the R2 of this model on the test set?

```{r}
y_pred = predict(yt_model_cart,yt_test)
y_true = yt_test$logviews
rss = sum((y_pred - y_true) ^ 2)
tss = sum((y_true - mean(y_true)) ^ 2)
1 - rss/tss
```


 
The out-of-sample R2 does not appear to be very good under regression trees, compared to a linear regression model. We could potentially improve it via cross validation.

Set seed to 100, run a 10-fold cross-validated cart model, with cp ranging from 0.0001 to 0.005 in increments of 0.0001. What is the optimal cp value on this grid?
```{r}
set.seed(100)
numFolds = trainControl(method = "cv", number = 10 )
cartGrid = expand.grid( .cp = seq(0.0001,0.005,0.0001))
train(logviews ~ logdislikes, data = yt_train, method = "rpart", trControl = numFolds, tuneGrid = cartGrid )
```


What is the R2 of this new model on the test set?
```{r}
yt_model_cart_2 = rpart(logviews ~ logdislikes,yt_train, cp=0.0001)
y_pred = predict(yt_model_cart_2,yt_test)
y_true = yt_test$logviews
rss = sum((y_pred - y_true) ^ 2)
tss = sum((y_true - mean(y_true)) ^ 2)
1 - rss/tss
```

Create a random forest model that predicts log(views) using the same variable as the CART model, with nodesize = 200 and ntree = 50. Set the random seed to 100.

```{r}
set.seed(100)
yt_model_rf = randomForest(logviews ~ logdislikes, yt_train, nodesize=200,ntree=50)
```

What is the R2 of this new model on the test set?
```{r}
y_pred = predict(yt_model_rf,yt_test)
y_true = yt_test$logviews
rss = sum((y_pred - y_true) ^ 2)
tss = sum((y_true - mean(y_true)) ^ 2)
1 - rss/tss
```


